### 2020-05-27 text/article_vedrovice.Rmd
text - permutation

The material analyzed here comes from Vedrovice – Široká u lesa cemetery.
Originally, the burial ground consisted of at least 108 burials.
Some of the burials were disturbed by later settlement and burial pits in the
area of the cemetery, amateur excavations during 19th century or later
agricultural activities.
Thus some of the burials are excluded from the analysis due to missing and/or
incomplete data and only a sample of 78 well-preserved burials is analyzed here.
Most of the data was obtained from the original publication [@podborsky2002],
anthropological determination of sex and age were corrected based on
more recent analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.

Each burial was originally characterized by 55 variables but only 35 are
selected for further analysis.
Some of  the original variables are omitted due to their rare occurrence,
for instance spondylus arm ring is present in a single burial only,
thus being useless in establishing relationships between pairs of burials.
Other variables are removed because they are highly dependent one on each other.

@@

Second large category of variables consists of data on artifacts present in the
burials, the grave goods.
These include variables noting quantities of various types of pottery,
positions of the pottery vessels within the graves, lithics and most common
raw materials of lithics, polished stone axes or adzes, grinding slabs,
pebbles, bone tools and different kinds of personal ornaments including beads
made of spondylus shell, pendants or oval shaped spondylus buckles and presence
of ochre pigment.
All the variables in this category are simplified by binarization (presence vs
absence of a trait) in order to minimize effects of only several numerous
cases on the overall pattern, for instance 20 beads in a grave <!-- 70 -->
would have much larger effect then 1 pottery vessels in the same grave.

@@

Data is analyzed in R environment for statistical computing and graphics
[@rcore]  using functions of readr [@wickham2017a], dplyr [@wickham2016] and
reshape2 [@wickham2007] packages to manipulate the data.
Function daisy [@kaufman1990] from package cluster [@maechler2016] is used to
count distances between observations used in further analyses.

>>>>

The material analyzed here comes from Vedrovice – Široká u lesa cemetery.
Originally, the burial ground consisted of at least 108 burials.
Some of the burials were disturbed by later settlement and burial pits in the
area of the cemetery, amateur excavations during the 19th century or later
agricultural activities.
Thus some of the burials are excluded from the analysis due to missing and/or
incomplete data and only a sample of 78 well-preserved burials is analyzed here.
Most of the data was obtained from the original publication [@podborsky2002],
anthropological determination of sex and age were corrected based on
more recent analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.

Each burial was originally characterized by 55 variables but only 35 are
selected for further analysis.
Some of  the original variables are omitted due to their rare occurrence,
for instance spondylus arm ring is present in a single burial only,
thus being useless in establishing relationships between pairs of burials.
Other variables are removed because they are highly dependent on each other.

@@

Second large category of variables consists of data on artifacts present in the
burials.
These include variables noting quantities of various types of pottery and
positions of the pottery vessels within the graves,
counts of various raw materials of chipped stone tools,
presence of polished stone axes or adzes, grinding slabs, pebbles,
bone tools and different kinds of personal ornaments.
The personal ornaments include beads made of spondylus shell or marble,
sponylus pendants or oval shaped buckles.
Presence of ochre pigment is also noted.
All the variables in this category are simplified by binarization (presence vs
absence of a trait) in order to minimize effects of only several numerous
cases on the overall pattern, for instance 20 beads in a grave <!-- 70 -->
would have much larger effect on the analysis then 1 pottery vessel in the same grave.

@@

Data is analyzed in an R environment for statistical computing and graphics
[@rcore]  using functions of readr [@wickham2017a], dplyr [@wickham2016] and
reshape2 [@wickham2007] packages to manipulate the data.
Function daisy [@kaufman1990] from package cluster [@maechler2016] is used to
count distances between observations used in further analyses.
<!-- ADD METHODS AND PACKAGES USED IN R -->

## Randomness of Artefact Combinations

To determine the non-randomness of various artefact combinations we conduct a
permutation test on the presence/absence data matrix.
The method based on a theory of generalized Monte Carlo tests was first
introduced to archaeology by @manly1996.
Original (observed) co-occurence matrix is permutated $n$ times with constant
row and column totals, i. e. both the count of artifacts in individual graves
and sum of the given type of artifacts are fixed.

To examine which associations of variables are infrequent or even not present
and which tend to appear together we conduct a randomization test
